Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-02-11 MCQMC2022ChallengesQMCSoftwareArticle.tex
2023 Mac 14 Updates

\subsection{LD Generators} A lattice, digital net, or Halton point generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.  These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.

LD generators are the most prevalent generally available components of QMC software.  These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.  Joe and Kuo

The \emph{integrand}, $g$, and the non-negative weight, $\lambda$, define the original integral arising from an application.  The function $g$ may represent the discounted payoff of a financial derivative, or the velocity of flow at a point in rock with stochastic porosity.  If $\lambda$ is a probability density function for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.

@@

LatNetBuilder \cite{LatNet}
   and MatBuilder \cite{paulin2022}
    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
    \item Multi-level software \cite{GilesSoft,GilesQSoft}
    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
ago
    \item CUBA \cite{CUBA}

>>>>

\subsection{LD Sequence Generators} A lattice, digital sequence, or Halton sequence generator typically supplies the sequence $\bsx_1, \bsx_2, \ldots$.  These sequences may be deterministic or random and are intended to have an empirical distribution that approximates well the uniform distribution.

LD generators are the most prevalent generally available components of QMC software.  These generators appear in Association of Computing Machinery  publications \cite{BraFox88,BraFoxNie92,HonHic00a}, FinDer \cite{PasTra95,FinDer}, libseq \cite{FriKel02,FriKelweb}, BRODA \cite{BRODA20a}, NAG \cite{NAG27}, MATLAB \cite{MAT9.13}, SamplePack \cite{SamplePack}, Gr\"nschlo{\ss}'s website \cite{GruWeb}, the Magic Point Shop \cite{NuyWeb}, R \cite{QRNG2020}, Julia \cite{QMCJulia}, SciPy \cite{virtanen2020scipy}, PyTorch \cite{paszke2019pytorch}, and TensorFlow \cite{tfqfQMC2021a}, and MatBuilder \cite{paulin2022}.

Lattice and digital sequences are not unique, and improved or alternative versions continue to be constructed.  Joe and Kuo \cite{JoeKuo03,JoeKuo08,SobolDirection} have proposed  Sobol' generator direction numbers, which are now widely used, but were not part of the early LD sequence software.  L'Ecuyer and colleagues have created LatNetBuilder \cite{LatNet} to construct good lattice and polynomial lattice sequences based on user defined criteria. As was raised during a lunch at MCQMC 2022, we should agree on a consistent format for storing the parameters that define good LD sequences so that they can by shared across software libraries.

\subsection{Integrands and Variable Transformations} The original integral or expectation arising from the application is defined in terms of the integrand, $g$, and the non-negative weight, $\lambda$.  For example,
\begin{itemize}
	\item $g$ may represent the discounted payoff of a financial derivative and $\lambda$ be the probability density function (PDF) for a discretized Brownian motion \cite{Gla03},

	\item  $g$ may represent the velocity of flow at a point in rock and $\lambda$ is the PDF of the porosity field , or

	\item $g\lambda$ is the unnormalized Bayesian posterior density multiplied by parameter of interest.
\end{itemize},
If $\lambda$ is a probability density function for a random variable $\bsT$, then $\mu = \EE[g(\bsT)]$.

@@

\iffalse
LatNetBuilder \cite{LatNet}
   and MatBuilder \cite{paulin2022}
    \item Sobol' direction numbers \cite{JoeKuo03,JoeKuo08,SobolDirection}
    \item Fast CBC, Magic Point Shop, QMC4PDE and other code since 2004  at \cite{NuyWeb}
    \item Multi-level software \cite{GilesSoft,GilesQSoft}
    \item Data-driven error bounds and stopping criteria in GAIL \cite{ChoEtal21a} and QMCPy \cite{QMCPy2020a, ChoEtal22a}
    \item Uncertainty quantification libraries Dakota \cite{DakotaUsersManual}, UQTk \cite{DebEtal04,UQTk}, and MUQ \cite{MUQ}  have some basic level low discrepancy sampling
ago
    \item CUBA \cite{CUBA}
\fi
